---
title: 'Reproducible Research: Peer Assessment 1'
output:
    html_document
keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
``` 


## Loading and preprocessing the data
1. Unzip the datas in the working directory

```{r}
unzip("activity.zip")
```

2. Read the datas

```{r}
data <- read.csv("activity.csv")
```


3. Have a look on the datas
Dimensions :
```{r}
dim(data)
```

We have 17568 rows and 3 colums, let look at their structure
```{r}
str(data)
```

And their first rows and the summary :

```{r}
head(data)
summary(data)
```

As we can see, the date are not interpreted as data, let's correct that

```{r}
data$date = as.Date(data$date)
summary(data)
```


## What is mean total number of steps taken per day?


1. Aggregage the steps per date and plot the histogram of steps taken per day
```{r}
steps_taken_day <- aggregate(steps ~ date, data=data, FUN=sum, na.RM=TRUE)
hist(steps_taken_day$steps, xlab = "Steps", main="Histogram of Steps taken per day")
```

2. Mean and median of steps taken per day

```{r}
steps_taken_day_mean <- mean(steps_taken_day$steps)
steps_taken_day_mean
steps_taken_day_median <- median(steps_taken_day$steps)
steps_taken_day_median
```

3. Report it, mean in red and median in green

```{r}
hist(steps_taken_day$steps, xlab = "Steps", main="Histogram of Steps taken per day")
abline(v=steps_taken_day_mean, col="red")
abline(v=steps_taken_day_median, col="green")

```


## What is the average daily activity pattern?


1. Aggregate the steps per interval and plot the average daily activity pattern
```{r}
daily_activity <- aggregate(steps ~ interval, data=data, FUN=mean, na.rm=TRUE)
plot(daily_activity, type="l", xlab="Interval", ylab="Steps")
```

2. 5-min interval containing maximum steps
```{r}
daily_activity[which.max(daily_activity$steps), ]
```

3. Report it
```{r}
plot(daily_activity, type="l", xlab="Interval", ylab="Steps")
abline(v=daily_activity[which.max(daily_activity$steps), ]$interval, col="red")
```

## Imputing missing values

1. Calculate and report the total number of missing values
```{r}
sum(is.na(data$steps))
```


2 and 3 : Create a new dataset with mising value replaced by the daily activity
for this interval

```{r}
complete_data <- data
complete_data[is.na(data$steps),]$steps <-
    sapply(data[is.na(data$steps), ]$interval,
           function(x){
               daily_activity[daily_activity$interval == x,]$steps
               }
           )
```

4. Make the histogram of the total of steps taken each day

* First, aggregate the data
* Compute the mean and the median
* Plot the histogram and report the mean (in red) and the median (in green)
* Report the previous mean (in orange) and the previous median (in purple)
in order to compare

```{r}
step_by_day <- aggregate(steps ~ date, data=complete_data, FUN=sum)
step_by_day_mean <- mean(step_by_day$steps)
step_by_day_median <- median(step_by_day$steps)
hist(step_by_day$steps, xlab="Steps", main="Histogram of steps by day")
abline(v=step_by_day_mean, col="red")
abline(v=steps_taken_day_mean, col="orange")
abline(v=step_by_day_median, col="green")
abline(v=steps_taken_day_median, col="purple")
```

There is only a purple line, this is due to the fact that the meadians and means
are close


```{r}
step_by_day_mean
steps_taken_day_mean
step_by_day_median
steps_taken_day_median
```

## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable with two levels : weekdays and weekend
(For reproductibility using POSIXlt : sunday are 0 and saturday 6)

```{r}
complete_data$day_type <- factor(sapply(
    complete_data$date,
    function(x) {
        if(as.POSIXlt(x)$wday %in% c(6, 0))
            return("weekend")
        else
            return("weekday")
        }
))
```

2. Compute the mean of weekday and weekend

```{r}
day_type_interval_mean <- aggregate(steps ~ interval + day_type,
                                    data=complete_data, FUN=mean)
```


2. Panel plot of 5 minutes intervals for weekend and weekday

```{r}
par(mfrow = c(2, 1))
plot(steps ~ interval, 
     data=day_type_interval_mean[day_type_interval_mean$day_type == "weekend",],
     type="l",
     main="Weekend",
     xlab="Interval",
     ylab="Steps")

plot(steps ~ interval, 
     data=day_type_interval_mean[day_type_interval_mean$day_type == "weekday",],
     type="l",
     main="Weekday",
     xlab="Interval",
     ylab="Steps")
```


